Bow shocks, bow waves, and dust waves. III. Diagnostics

### 2019-06-21 papers/Papers1234/bs-bw-dw-03.tex
Finish addressing point 2 of ref report for paper 3

\bibliographystyle{mnras}
\bibliography{bowshocks-biblio}
\appendix
\input{app-k17-k18}

>>>>

\bibliographystyle{mnras}
\bibliography{bowshocks-biblio}
\appendix

\section{Dust opacity and optical depth}
\label{app:dust-opacity}

In the main body of the paper we use a frequency-averaged dust optical
depth, \(\tau\), which we here relate to the frequency-dependent optical
properties of the grains.  For simplicity, we assume a dust model that
consists of a set \(k \in \{1 \dots N\}\) of discrete populations of
solid spherical grains \citep[e.g.][]{Desert:1990a}, each with its own
chemical composition and size distribution:
\(y_{a,k} \equiv n_k^{-1} (d n_k / d a)\), where \(n_k\) is the volume
density of the \(k\)-th population (\si{grain.cm^{-3}}) and \(a\) is
the grain radius. The frequency-dependent dust opacity per unit mass
of gas and dust is \(\kappa_\nu\), with units \si{cm^2.g^{-1}}, which can be
expressed as
\begin{equation}
  \label{eq:dust-opacity-frequency}
  \kappa_\nu = \rho^{-1} \sum_{k=1}^{N} \left[  n_k \int_0^\infty \!\!y_{a,k} \, Q_{a,k,\nu} \,\pi a^2 \, da\right]\ ,
\end{equation}
where \(\rho\) is the total material mass density and \(Q_{a,k,\nu}\) is
the size-, composition-, and frequency-dependent dimensionless grain
optical efficiency, which can be calculated from the grain refractive
indices using Mie theory \citep{Bohren:1983a}.  There are different
varieties of \(Q\), and therefore different varieties of opacity,
according to whether one is interested in true absorption,
\(Q_{\text{abs}}\), scattering, \(Q_{\text{scat}}\), or total
extinction, \(Q_{\text{ext}} = Q_{\text{abs}} + Q_{\text{scat}}\).
The relative importance of scattering is described by the grain
albedo: \(\varpi = Q_{\text{scat}}/Q_{\text{ext}}\).  The radiation
pressure acting on the grains is determined by a fourth variety of
optical efficiency:
\(Q_P = Q_{\text{abs}} + (1 - g) Q_{\text{scat}}\), where \(g\) is the
mean scattering cosine.

The monochromatic optical depth of a medium (in any of these 4
varieties) along a path \(s\) is given by
\begin{equation}
  \label{eq:tau-monochromatic}
  \tau_\nu = \int \rho \kappa_\nu \, ds \ .
\end{equation}
If the effects of multiple scattering are ignored, then the fraction
of stellar radiation of frequency \(\nu\) that is absorbed by the shell
is \(1 - e^{-\tau_\nu}\), in which the integral is performed with
\(s\) along the radial direction, assuming the star to be effectively
a point source at the position of the shell.  To find the fraction of
radiative \emph{energy}, \(\kappa_\nu\) must be calculated using
\(Q_{\text{abs}}\), whereas to find the fraction of radiative
\emph{momentum}, it must be calculated using \(Q_P\).  This
single-scattering regime \citep[e.g.][]{Wibking:2018a} is a good
approximation so long as \(\varpi\) is not close to unity and the
shell is optically thin at infrared wavelengths, which is always true
for stellar bow shocks.

The frequency-averaged optical depth \(\tau\) is defined such that the
fraction of the frequency-integrated stellar radiative momentum
absorbed by the shell is \(1 - e^{-\tau}\).  For a stellar spectrum of
shape \(F_\nu\), this implies the formal relation
\begin{equation}
  \label{eq:tau-averaged}
  \tau = - \ln \left[ \frac{\int_0^\infty F_\nu \, e^{-\tau_\nu} \, d\nu}{\int_0^\infty F_\nu  \, d\nu} \right] \ .
\end{equation}
Note that this equation is valid regardless of the physical units of
\(F_\nu\) (e.g., flux or luminosity).  More approximately, \(\tau\) can
simply be estimated by \(\tau_\nu\) at the peak frequency of the stellar
radiation, which typically falls in the far-ultraviolet range,
although this approximation breaks down if \(\tau_\nu > 1\) over a
significant portion of the stellar spectrum.

In the main body of the paper, we use \(\tau\) for the radiation pressure
optical depth and \(\tau'\) for the absorption optical depth.  These are
related by the ratio of the corresponding frequency-averaged optical
efficiencies:
\begin{equation}
  \label{eq:tau_tauprime}
  \tau = \frac{Q_P}{Q_{\text{abs}}} \tau' = \left(  1 + \varpi \frac{1 - g}{1 - \varpi}\right) \tau'\ .
\end{equation}
Specific results were given in Figure~6a of Paper~II for an example
grain mixture, designed to reproduce average interstellar medium dust
properties \citep{Weingartner:2001b, Abel:2008a, Ferland:2013a,
  Ferland:2017a}.  For such a mixture the albedo is roughly constant
at \(\varpi \approx 0.5\) throughout the ultraviolet and visible bands,
falling to low values in the infrared.  The asymmetry parameter, or
mean scattering cosine, falls across the ultraviolet band from
\(g \approx 1\) at the highest frequencies to \(g\approx 0.5\) in the FUV.\@ This
yields a conversion factor of
\(Q_P / Q_{\text{abs}} = \text{\numrange{1.2}{1.3}}\) at EUV/FUV
wavelengths.

\input{app-k17-k18}

### 2019-06-21 papers/Papers1234/sec-observational-diagnostics.tex
Finish addressing point 2 of ref report for paper 3

A fundamental parameter is the optical depth, \(\tau\), of the bow
shell to UV radiation, which determines what fraction of the stellar
photon momentum is available to support the shell (see
\S~2.1 of Paper~I).  But the same photons also heat the
dust grains in the bow, which re-radiate that energy predominantly at
mid-infrared wavelengths (roughly \SIrange{10}{100}{\um}) with
luminosity \(L\IR\).  Assuming that Ly\(\alpha\) and mechanical
heating of the dust shell is negligible (see
\S~\ref{sec:unimp-other-heat}) and that the emitting shell subtends a
solid angle \(\Omega\), as seen from the star, then the optical depth
can be estimated as
\begin{equation}
  \label{eq:tau-empirical}
  \tau = -\ln \left( 1 - \frac{4\pi}{\Omega} \frac{L\IR}{L_*} \right)
  \approx \frac{2 L\IR}{L_*} \ ,
\end{equation}
where the last approximate equality holds if \(\tau \ll 1\) and the shell
emission covers one hemisphere.\footnote{%
  Note that the \(\tau\) of Paper~I is not exactly the same as the
  \(\tau\) of equation~\eqref{eq:tau-empirical}, but is larger by a
  factor of
  \(Q_P / Q_{\text{abs}} = 1 + \varpi (1 - g)/(1 - \varpi)\), where
  \(\varpi\) is the grain albedo and \(g\) the scattering asymmetry.
  For standard ISM grain mixtures,
  \(Q_P / Q_{\text{abs}} = \text{\numrange{1.2}{1.3}}\) at EUV/FUV
  wavelengths.}
% This is plotted in Figure~6a of Paper~II for Cloudy's default ISM
% grain mixture (see \S~\ref{sec:grain-temp-emiss} below), where it
% can be seen that
% \(Q_P / Q_{\text{abs}} = \text{\numrange{1.2}{1.3}}\) at EUV/FUV
% wavelengths (\(\lambda = \text{\SIrange{0.05}{0.2}{\um}}\)), so the
% correction is small.

>>>>

A fundamental parameter is the average optical depth, \(\tau\), of the
bow shell to the stellar radiation. In Appendix~\ref{app:dust-opacity}
we discuss in detail the relation between \(\tau\) and the optical
properties of the grains.  In the single-scattering approximation, a
fraction \(1 - e^{-\tau}\) of the stellar photon momentum is available to
support the shell (see \S~2.1 of Paper~I).  But the same photons also
heat the dust grains in the bow, which re-radiate that energy
predominantly at mid-infrared wavelengths (roughly
\SIrange{10}{100}{\um}) with luminosity \(L\IR\).  Assuming that
Ly\(\alpha\) and mechanical heating of the dust shell is negligible (see
\S~\ref{sec:unimp-other-heat}) and that the emitting shell subtends a
solid angle \(\Omega\), as seen from the star, then an absorption optical depth can
be estimated as
\begin{equation}
  \label{eq:tau-empirical}
  \tau' = -\ln \left( 1 - \frac{4\pi}{\Omega} \frac{L\IR}{L_*} \right)
  \approx \frac{2 L\IR}{L_*} \ ,
\end{equation}
where the last approximate equality holds if \(\tau' \ll 1\) and the shell
emission covers one hemisphere.  The two optical depths, \(\tau\) and
\(\tau'\), are not exactly the same, see
equation~\eqref{eq:tau_tauprime}.  The first, \(\tau\), represents
radiative momentum transfer, which is governed by the grain radiation
pressure efficiency, \(Q_P\).  The second, \(\tau'\), represents
radiative energy transfer, which is governed by the grain absorption
efficiency, \(Q_{\text{abs}}\).  We adopt \(\tau = 1.25 \tau'\) based on
Cloudy models of a standard ISM dust mixture, see Figure~6 of
Paper~II and Appendix~\ref{app:dust-opacity}.
